Introducing Statistical Design of Experiments to SPARQL Endpoint Evaluation

### 2013-05-04 exp1/iswc2013.tex
more on experimental setup
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@497 43ef1190-90fd-4dd7-a769-d0121f02b80e

%\documentclass{llncs}
\documentclass{article}
\usepackage{cite}
\usepackage{verbatim}

@@

As the focus of this paper is didactical, the actual measurements are
much simpler than what has been used in benchmarking.

@@

@@

We have used the dataset of \cite{mxro:Morsey2011DBpedia}, where

>>>>

%\documentclass{llncs}
\documentclass{article}
\usepackage{cite}
\usepackage{verbatim}
\newcommand{\todo}[1]{\ensuremath{^{\textrm{\tiny{TODO}\normalsize}}}\footnote{\textbf{TODO:}~#1}}

@@

As the focus of this paper is didactical, the actual measurements are
much simpler than what has been used in benchmarking. We have used the
dataset of \cite{mxro:Morsey2011DBpedia}, but only consider the
smallest of their datasets (that we found to me more than 15
MTriples). Moreover, we have taken subsets of that dataset by using
the last 1 or 2 MTriples their file.

@@

These fragments has been carefully designed for illustrative utility,
as well as suitable selectivity. When the experiment is run, they are
combined to yield a query, which is then sent to a SPARQL
endpoint. The SPARQL endpoint is itself set up using 4store\todo{ref}
version 1.1.5. The ``Machine'' factor specifies a hostname, whereas
``TripleC'' and ``Implement'' specifies a port number. On each host,
four 4store instances runs independently on different
ports. \todo{possible problem?}

@@

However, as a real optimization is beyond the scope of this paper, we
have instead created a simulation of an optimization. This has also
the advantage that we can \textit{a priori} understand the effect of
the changes. In a real-world case, there is often the case that an
optimization has negative side-effects, and we would need to simulate
both the optimization and the side-effects. To do that, we degraded
the performance of a the 4store SPARQL implementation by inserting
\texttt{sleep} statements. Specifically we inserted the C statement
\texttt{sleep(2)} on line~920 in \texttt{src/frontend/query.c} on
level~1, to simulate the optimization. This has the effect of delaying
execution for 2~seconds for every block for all kinds of joins. On
level 2, we inserted \texttt{usleep(2000)} on line 987 in
\texttt{src/frontend/filter.c} to simulate the negative
side-effect. This delays execution for 2 milliseconds every time the
\texttt{langMatches} SPARQL function is called.
